An efficient composition of bidirectional programs by tupling and lazy updates

### 2020-02-21 paper/1.introduction.tex
remove red color part & small fix

\noindent
\textcolor{red}{Unlike function composition, the composition of bidirectional programs is read left-to-right. We use this order because it is helpful to understand the behavior if we consider data goes from left to right.}
%\textcolor{red}{Note that the order of composition is the reverse order compared to function composition. We use this order, because it is standard in BX community.}
One feature of this composition is that $put_{bx_1 \tilde{\circ} bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficiency problem if we compute $put$ for composition of many bidirectional programs. Generally, for a composition of $O(n)$ bidirectional programs, we need to call $get$ for $O(n^2)$ times. To be concrete, consider the evaluation of the following composition (which will be used as our running example in this paper):

@@

To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
\textcolor{red}{Here, complements are information from sources that makes $get$ injective, which is in turn needed to evaluate $put$.
  In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead},
  we use $i_1$, $[[1,2],[3]]$ and $i_2'$, $[100,2]$ to obtain the updated source $[[100,2],[3]]$. However, $[1,2]$ is simply replaced by $[100,2]$ and not used to construct the result of $put$. In this case we can use $[..., [3]]$ as a complement.}
The key idea of the second approach is straightforward: Complements are smaller
% program fragments
than
% the original
intermediate states.
%Therefore this strategy solves the previous two problems.
%Readers might already know that it is a hard problem to obtain complements in general. In our work, thanks to the following finding, it is not hard:
%
%\vspace{2mm}
%In very-well behaved \cite{Foster:2007:CBT:1232420.1232424} BX,% programs,
%$put$ is a complement function for $get$.
%, $get$ can be a complement function for $put$.
%\vspace{2mm}
%
For obtaining complements, we tuple $put$ and $get$, and produce a new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
Let us reconsider our example $phead~\tilde{\circ}~phead~\tilde{\circ}~phead$ in Figure \ref{fig:eval-comp-phead-2}, where $c_1$ and $c_2$ are complements and $d_1$ and $d_2$ are valid views for $s$ and~$i_1$. Here, two points are worth noting. First, after evaluating the leftmost $pg$, the original source $s$ need not be kept because its complete contents are in $c_1$ and~$i_1$. Second, the complements are smaller than the intermediate states in Figure~\ref{fig:eval-comp-phead}.
%This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources.
Actually, this simple
% combined
$pg$ alone is not yet effective for left associative compositions because it requires two more $put$s, which can be seen on the right side of Figure \ref{fig:eval-comp-phead-2}. To achieve an efficient evaluation, we combine two techniques, lazy update and lazy evaluation. We explain this second method and all optimizations in Section~\ref{sect:xpg}.

>>>>

\noindent
Unlike function composition, the composition of bidirectional programs is read left-to-right. We use this order because it is helpful to understand the behavior if we consider data goes from left to right.
%\textcolor{red}{Note that the order of composition is the reverse order compared to function composition. We use this order, because it is standard in BX community.}
One feature of this composition is that $put_{bx_1 \tilde{\circ} bx_2}$ needs to call $get_{bx_1}$ to compute the intermediate result for $put_{bx_2}$ to use, which would introduce an efficiency problem if we compute $put$ for composition of many bidirectional programs. Generally, for a composition of $O(n)$ bidirectional programs, we need to call $get$ for $O(n^2)$ times. To be concrete, consider the evaluation of the following composition (which will be used as our running example in this paper):

@@

To treat large inputs, we propose the second method based on memoization: ``keeping \emph{complements in a closure} and using them when needed''.
Here, complements are information from sources that makes $get$ injective, which is in turn needed to evaluate $put$.
  In the middle $put$ and $get$ of Figure \ref{fig:eval-comp-phead},
  we use $i_1$, $[[1,2],[3]]$ and $i_2'$, $[100,2]$ to obtain the updated source $[[100,2],[3]]$. However, $[1,2]$ is simply replaced by $[100,2]$ and not used to construct the result of $put$. In this case we can use $[..., [3]]$ as a complement.
The key idea of the second approach is straightforward: Complements are smaller
% program fragments
than
% the original
intermediate states.
%Therefore this strategy solves the previous two problems.
%Readers might already know that it is a hard problem to obtain complements in general. In our work, thanks to the following finding, it is not hard:
%
%\vspace{2mm}
%In very-well behaved \cite{Foster:2007:CBT:1232420.1232424} BX,% programs,
%$put$ is a complement function for $get$.
%, $get$ can be a complement function for $put$.
%\vspace{2mm}
%
For obtaining complements, we tuple $put$ and $get$, and produce a new function $pg$. Because $put$ produces new complements for $get$, we can shrink the size.
Let us reconsider our example $phead~\tilde{\circ}~phead~\tilde{\circ}~phead$ in Figure \ref{fig:eval-comp-phead-2}, where $c_1$ and $c_2$ are complements and $d_1$ and $d_2$ are valid views for $s$ and~$i_1$. Here, two points are worth noting. First, after evaluating the leftmost $pg$, the original source $s$ need not be kept because its complete contents are in $c_1$ and~$i_1$. Second, the complements are smaller than the intermediate states in Figure~\ref{fig:eval-comp-phead}.
%This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources.
Actually, this simple
% combined
$pg$ alone is not yet effective for left associative compositions because it requires two more $put$s, which can be seen on the right side of Figure \ref{fig:eval-comp-phead-2}. To achieve an efficient evaluation, we combine two techniques, lazy update and lazy evaluation. We explain this second method and all optimizations in Section~\ref{sect:xpg}.

### 2020-02-21 paper/2.minbigul.tex
remove red color part & small fix

To help make demonstration more direct, we provide the following alternatives representation: $Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2, \ Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$. \textcolor{blue}{The compose symbol $\tilde{\circ}$ used in the previous section will be replaced with the more common one, $\circ$.}
In general, $\circ$ has a higher priority than $\times$. Their associativity precedence can be either left or right or mixture, but are not set by default. We need to explicitly write programs that use these operators.

>>>>

To help make demonstration more direct, we provide the following alternatives representation: $Prod \ bx_1 \ bx_2 \equiv bx_1 \times bx_2, \ Compose \ bx_1 \ bx_2 \equiv bx_1 \circ bx_2$. The compose symbol $\tilde{\circ}$ used in the previous section will be replaced with the more common one, $\circ$.
In general, $\circ$ has a higher priority than $\times$. Their associativity precedence can be either left or right or mixture, but are not set by default. We need to explicitly write programs that use these operators.

### 2020-02-21 paper/4.xpg.tex
small changes: \circ -> \tilde{\circ} in intro. sec., \smallvspace
remove red color part & small fix

    \noindent $\tabs{2} \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{\textit{undefined}}, h \ s) \\
        \tab \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{\textit{undefined}} \\
        \tab \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{\textit{undefined}}$

@@

    \noindent $\tabs{2} \pg{Replace}{s}{v} = (v, s)$\\
    $\tabs{2} \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
        \tab \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1},\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}, \getbx{bx_2}{s_2})) \\
        \tab \qtab \overset{2}{=} (s_1', v_1') \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
            \qtab \tabs{2} \ \, (s_2', v_2') \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
            \tab \qtabs{2} ((s_1',s_2'), (v_1',v_2'))$\\
    $\tabs{2} \pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \overset{1}{=} (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
        \tab \qtab \overset{2}{=} (s', v') \Leftarrow \pg{bx}{f_1 \ s}{v};\\
            \tab \qtabs{2} (f_2 \ s', v')$\\
    $\tabs{2} \pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \overset{1}{=} (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
        \tab \qtab \overset{2}{=} (s', v') \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
        \tab \qtabs{2} (s', g_2 \ v')$

@@

    \noindent $\tabs{2} \pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
    \tabs{3} \overset{1}{=} (\text{if} \ {cond_{sv} \ s \ v} \qtabs{3} \tab \text{if} \ {cond_{s} \ s} \\
    \tabs{2} \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \qtab \tab \, \text{then} \ v' \Leftarrow \getbx{bx_1}{s} \\
    \tabs{2} \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \qtab \tab \ \, \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
    \tabs{2} \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s' \qtab , \tab \ \ \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v')\\
    %
    \tabs{3} \overset{2}{=} \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \ \text{then}\\
        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_1}{s}{v}, \getbx{bx_1}{s});\\
        \qtabs{2} \tab \text{if} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v' \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
        \tabs{2} \qtab \text{else if} \ {cond_{sv} \ s \ v} \ \&\& \ not \ ({cond_{s} \ s}) \ \text{then}\\
        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_1}{s}{v}, \getbx{bx_2}{s});\\
        \qtabs{2} \tab \text{if} \ cond_{s} \ s' \ \&\& \ not \ (cond_{sv} \ s \ v') \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
        \tabs{2} \qtab \text{else if} \ not \ ({cond_{sv} \ s \ v}) \ \&\& \ {cond_{s} \ s} \ \text{then}\\
        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_2}{s}{v}, \getbx{bx_1}{s});\\
        \qtabs{2} \tab \text{if} \ not \ (cond_{s} \ s') \ \&\& \ cond_{sv} \ s \ v' \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
        \tabs{2} \qtab \text{else if} \ not \ ({cond_{sv} \ s \ v}) \ \&\& \ not \ ({cond_{s} \ s}) \ \text{then}\\
        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_2}{s}{v}, \getbx{bx_2}{s});\\
        \qtabs{2} \tab \text{if} \ not \ (cond_{s} \ s') \ \&\& \ not \ (cond_{sv} \ s \ v') \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
    %
    \tabs{3} \overset{3}{=} (* \text{ with restriction } *)\\
    \tabs{2} \qtab \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s}\\
    \tabs{2} \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
    \tabs{2} \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
    \tabs{2} \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$

@@

    \noindent $\tabs{2} \pg{bx_1 \circ bx_2}{s}{v} \\
    \tabs{3} \overset{1}{=} (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
    \tabs{3} \overset{2}{=} v_1 \Leftarrow \getbx{bx_1}{s}; \qtabs{3} \ \, \overset{3}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy};\\
        \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \qtabs{2} (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v};\\
        \tabs{2} \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \qtabs{2} (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2};\\
            \qtab \tabs{3} (s_3, v_2) \qtabs{6} (s_3, v_2)\\
    \tabs{3} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
        \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
        \tabs{2} \qtab (s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
            \qtab \tabs{3} (s_3, v_2)$

@@

$\tab \cpg{bx}{ks}{kv}{s}{v} = \pg{bx}{ks \ s}{kv \ v}$

@@

$\tab \kpg{bx}{ks}{kv}{lks}{lkv}{s}{v} = \cpg{bx}{ks \circ lks}{kv \circ lkv}{s}{v}\\
    \tabs{5} = \pg{bx}{ks \ (lks \ s)}{kv \ (lkv \ v)}$

@@

\textcolor{red}{
In $kpg$, basically, functions for the updates are kept (but not evaluated) in $lks$ and $lkv$.
In $\kpginline{RearrS}$ and $\kpginline{RearrV}$, $f_1$ and $g_1$ are accumulated in $lks$ and $lkv$.
The kept functions are evaluated in $\kpginline{Skip}$ and $\kpginline{Case}$ by applications of $lks \ s$ and $lkv \ v$. At the same time, the third and fourth argument of recursive calls are updated with the identity function.
This evaluation is needed because these definitions require the actual values, $es$ and $ev$.
Thanks to this update, accumulation in $kpg$, $\underline{lks_1}$ and $\underline{s_1}$ in $\kpginline{bx_1 \circ bx_2}$ are not evaluated as much as possible.}
%Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.}

\textcolor{red}{Additionally we did two optimizations in $kpg$. The first is in $\kpginline{bx_1 \times bx_2}$.
  Because $es$ and $ev$ are not used in this definition, we do not need to evaluate.
  However, if we accumulate $lks$ and $lkv$, both might be evaluated independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$. This includes the same computation. To remove duplicate evaluations, we evaluate actual values $es$ and $ev$ before calling $\kpginline{bx_1}$. and $\kpginline{bx_2}$}.
%It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that. One more thing,
The second is in $\kpginline{Case}$ and not shown in the definition. We need to evaluate $lks' \ s'$ and $lkv' \ v'$ to check the \underline{fi} condition before returning the 6-tuple. Such evaluations can be done lazily to make programs run faster. We use the above small optimizations in our implementation.

>>>>

\smallvspace
    \noindent $\tabs{2} \pg{Skip \ h}{s}{v} \overset{1}{=} (\sif{h \ s = v}{s}{\textit{undefined}}, h \ s) \\
        \tab \qtab \overset{2}{=} \sif{h \ s = v}{(s, h \ s)}{\textit{undefined}} \\
        \tab \qtab \overset{3}{=} \sif{h \ s = v}{(s, v)}{\textit{undefined}}$
\smallvspace

@@

\smallvspace
    \noindent $\tabs{2} \pg{Replace}{s}{v} = (v, s)$\\
    $\tabs{2} \pg{\product{bx_1}{bx_2}}{(s_1,s_2)}{(v_1,v_2)}\\
        \tab \qtab \overset{1}{=} ((\putbx{bx_1}{s_1}{v_1},\putbx{bx_2}{s_2}{v_2}), (\getbx{bx_1}{s_1}, \getbx{bx_2}{s_2})) \\
        \tab \qtab \overset{2}{=} (s_1', v_1') \Leftarrow \pg{bx_1}{s_1}{v_1}; \\
            \qtab \tabs{2} \ \, (s_2', v_2') \Leftarrow \pg{bx_2}{s_2}{v_2}; \\
            \tab \qtabs{2} ((s_1',s_2'), (v_1',v_2'))$\\
    $\tabs{2} \pg{\rearrs{f_1}{f_2}{bx}}{s}{v} \overset{1}{=} (f_2 \ (\putbx{bx}{(f_1 \ s)}{v}), \getbx{bx}{(f_1 \ s)}) \\
        \tab \qtab \overset{2}{=} (s', v') \Leftarrow \pg{bx}{f_1 \ s}{v};\\
            \tab \qtabs{2} (f_2 \ s', v')$\\
    $\tabs{2} \pg{\rearrv{g_1}{g_2}{bx}}{s}{v} \overset{1}{=} (\putbx{bx}{s}{(g_1 \ v)}, g_2 \ (\getbx{bx}{s})) \\
        \tab \qtab \overset{2}{=} (s', v') \Leftarrow \pg{bx}{s}{g_1 \ v}; \\
        \tab \qtabs{2} (s', g_2 \ v')$
\smallvspace

@@

\smallvspace
    \noindent $\tabs{2} \pg{\casebx{cond_{sv}}{cond_{s}}{bx_1}{bx_2}}{s}{v} \\
    \tabs{3} \overset{1}{=} (\text{if} \ {cond_{sv} \ s \ v} \qtabs{3} \tab \text{if} \ {cond_{s} \ s} \\
    \tabs{2} \qtab \text{then} \ s' \Leftarrow \putbx{bx_1}{s}{v} \qtab \tab \, \text{then} \ v' \Leftarrow \getbx{bx_1}{s} \\
    \tabs{2} \qtab \text{else} \ s' \Leftarrow \putbx{bx_2}{s}{v} \qtab \tab \ \, \text{else} \ v' \Leftarrow \getbx{bx_2}{s}\\
    \tabs{2} \qtab \text{fi} \ cond_{s} \ s'; \text{ return } s' \qtab , \tab \ \ \text{fi} \ {cond_{sv} \ s \ v'}; \text{ return } v')\\
    %
    \tabs{3} \overset{2}{=} \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s} \ \text{then}\\
        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_1}{s}{v}, \getbx{bx_1}{s});\\
        \qtabs{2} \tab \text{if} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v' \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
        \tabs{2} \qtab \text{else if} \ {cond_{sv} \ s \ v} \ \&\& \ not \ ({cond_{s} \ s}) \ \text{then}\\
        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_1}{s}{v}, \getbx{bx_2}{s});\\
        \qtabs{2} \tab \text{if} \ cond_{s} \ s' \ \&\& \ not \ (cond_{sv} \ s \ v') \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
        \tabs{2} \qtab \text{else if} \ not \ ({cond_{sv} \ s \ v}) \ \&\& \ {cond_{s} \ s} \ \text{then}\\
        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_2}{s}{v}, \getbx{bx_1}{s});\\
        \qtabs{2} \tab \text{if} \ not \ (cond_{s} \ s') \ \&\& \ cond_{sv} \ s \ v' \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
        \tabs{2} \qtab \text{else if} \ not \ ({cond_{sv} \ s \ v}) \ \&\& \ not \ ({cond_{s} \ s}) \ \text{then}\\
        \qtabs{2} \tab (s', v') \Leftarrow (\putbx{bx_2}{s}{v}, \getbx{bx_2}{s});\\
        \qtabs{2} \tab \text{if} \ not \ (cond_{s} \ s') \ \&\& \ not \ (cond_{sv} \ s \ v') \ \text{then return } (s',v') \ \text{else } \textit{undefined}\\
    %
    \tabs{3} \overset{3}{=} (* \text{ with restriction } *)\\
    \tabs{2} \qtab \text{if} \ {cond_{sv} \ s \ v} \ \&\& \ {cond_{s} \ s}\\
    \tabs{2} \qtab \text{then} \ (s', v') \Leftarrow \pg{bx_1}{s}{v}\\
    \tabs{2} \qtab \text{else} \ (s', v') \Leftarrow \pg{bx_2}{s}{v}\\
    \tabs{2} \qtab \text{fi} \ cond_{s} \ s' \ \&\& \ cond_{sv} \ s \ v'; \text{ return } (s',v')$
\smallvspace

@@

\smallvspace
    \noindent $\tabs{2} \pg{bx_1 \circ bx_2}{s}{v} \\
    \tabs{3} \overset{1}{=} (\putbx{bx_1}{s}{(\putbx{bx_2}{(\getbx{bx_1}{s})}{v})}, \getbx{bx_2}{(\getbx{bx_1}{s})}) \\
    \tabs{3} \overset{2}{=} v_1 \Leftarrow \getbx{bx_1}{s}; \qtabs{3} \ \, \overset{3}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy};\\
        \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \qtabs{2} (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v};\\
        \tabs{2} \qtab (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2}; \qtabs{2} (s_3, v_3) \Leftarrow \pg{bx_1}{s}{s_2};\\
            \qtab \tabs{3} (s_3, v_2) \qtabs{6} (s_3, v_2)\\
    \tabs{3} \overset{4}{=} (s_1, v_1) \Leftarrow \pg{bx_1}{s}{dummy}; \\
        \tabs{2} \qtab (s_2, v_2) \Leftarrow \pg{bx_2}{v_1}{v}; \\
        \tabs{2} \qtab (s_3, v_3') \Leftarrow \pg{bx_1}{s_1}{s_2}; \\
            \qtab \tabs{3} (s_3, v_2)$
\smallvspace

@@

\smallvspace
$\tab \cpg{bx}{ks}{kv}{s}{v} = \pg{bx}{ks \ s}{kv \ v}$
\smallvspace

@@

\smallvspace
$\tab \kpg{bx}{ks}{kv}{lks}{lkv}{s}{v} = \cpg{bx}{ks \circ lks}{kv \circ lkv}{s}{v}\\
    \tabs{5} = \pg{bx}{ks \ (lks \ s)}{kv \ (lkv \ v)}$
\smallvspace

@@

In $kpg$, basically, functions for the updates are kept (but not evaluated) in $lks$ and $lkv$.
In $\kpginline{RearrS}$ and $\kpginline{RearrV}$, $f_1$ and $g_1$ are accumulated in $lks$ and $lkv$.
The kept functions are evaluated in $\kpginline{Skip}$ and $\kpginline{Case}$ by applications of $lks \ s$ and $lkv \ v$. At the same time, the third and fourth argument of recursive calls are updated with the identity function.
This evaluation is needed because these definitions require the actual values, $es$ and $ev$.
Thanks to this update, accumulation in $kpg$, $\underline{lks_1}$ and $\underline{s_1}$ in $\kpginline{bx_1 \circ bx_2}$ are not evaluated as much as possible.
%Except $\kpginline{Skip}$ and $\kpginline{Case}$ where $s$ and $v$ hold actual values, functions for the computation will be kept in $ks'$ and $kv'$.}

Additionally we did two optimizations in $kpg$. The first is in $\kpginline{bx_1 \times bx_2}$.
  Because $es$ and $ev$ are not used in this definition, we do not need to evaluate.
  However, if we accumulate $lks$ and $lkv$, both might be evaluated independently in two assignments using $\kpginline{bx_1}$ and $\kpginline{bx_2}$. This includes the same computation. To remove duplicate evaluations, we evaluate actual values $es$ and $ev$ before calling $\kpginline{bx_1}$ and $\kpginline{bx_2}$.
%It is possible to evaluate actual values in $s$ and $v$ before calling $\kpginline{bx_1}$ to remove the redundancy like that. One more thing,
The second is in $\kpginline{Case}$ and not shown in the definition. We need to evaluate $lks' \ s'$ and $lkv' \ v'$ to check the \underline{fi} condition before returning the 6-tuple. Such evaluations can be done lazily to make programs run faster. We use the above small optimizations in our implementation.

### 2020-02-21 paper/5.experiments.tex
remove red color part & small fix

We have fully implemented and tested all methods% 
%
\footnote{All experiments on macOS 10.14.6, processor Intel Core i7 (2.6~GHz), RAM 16~GB 2400~MHz DDR4, OCaml 4.07.1. The OCaml runtime system options and garbage collection parameters are set as default.}$^,$%HACK!
%
\footnote{\textcolor{red}{The implementation is available:} \url{https://github.com/k-tsushima/pgs}}
%
described in the previous sections. Our target language is untyped. Some dummies used for $pg$, $cpg$, $kpg$ and $xpg$ are replaced with the current updated views. This helps a program in the $put$ direction valid.

@@

\textcolor{red}{Note that this result concerns BX programs that uses many compositions. If the number of compositions is small, $put$ will be fastest because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).}

>>>>

We have fully implemented and tested all methods% 
%
\footnote{All experiments on macOS 10.14.6, processor Intel Core i7 (2.6~GHz), RAM 16~GB 2400~MHz DDR4, OCaml 4.07.1. The OCaml runtime system options and garbage collection parameters are set as default.}$^,$%HACK!
%
\footnote{The implementation is available: \url{https://github.com/k-tsushima/pgs}}
%
described in the previous sections. Our target language is untyped. Some dummies used for $pg$, $cpg$, $kpg$ and $xpg$ are replaced with the current updated views. This helps a program in the $put$ direction valid.

@@

Note that this result concerns BX programs that uses many compositions. If the number of compositions is small, the original $put$ without any memoization will be fastest because of the overhead for memoization ($put_m$) and the overhead for keeping complements in closures ($xpg$).

### 2020-02-21 paper/6.related-work.tex
remove red color part & small fix

Since the pioneering work of lens \cite{Foster:2007:CBT:1232420.1232424}, many BX languages have been proposed \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487,Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}.
%(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
%based on several application scenarios and BX lenses works \cite{}.
Although much progress has been made on the semantics and correctness of BX programs for the past years, as far as we are aware, little work has been done on optimization of BX programs \cite{Horn:2018:IRL:3243631.3236769}. Anjorin et. al. introduces the first benchmark for BX languages and compared them \cite{Anjorin2019}, but a systematic improvement for practical implementation of BX languages is still missing. This paper shows the first attempt of improving efficiency of BX composition evaluation.

>>>>

Since the pioneering work of lens \cite{Foster:2007:CBT:1232420.1232424}, many BX languages have been proposed \cite{Bohannon06relationallenses:, Bohannon:2008:BRL:1328438.1328487,Buchmann:2018:BFI:3362232.3362263,Ko:2016:BFV:2847538.2847544,LeblebiciAS14,Samimi-Dehkordi18,Cicchetti2011,Hinkel:2019:CPB:3318595.3318634}.
%(memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?)
%based on several application scenarios and BX lenses works \cite{}.
Although much progress has been made on the semantics and correctness of BX programs for the past years, as far as we are aware, little work has been done on optimization of BX programs \cite{Horn:2018:IRL:3243631.3236769}. Anjorin et. al. introduces the first benchmark\footnote{The BX programs in their benchmark are basically BX programs without composition. Because we focus on the BX programs that include many compositions, their benchmark is not applicable for our evaluation.} for BX languages and compared them \cite{Anjorin2019}, but a systematic improvement for practical implementation of BX languages is still missing. This paper shows the first attempt of improving efficiency of BX composition evaluation.

### 2020-02-21 paper/7.conclusion.tex
remove red color part & small fix

We will continue our work on the following three points.
First is overcoming our limitation in $xpg$. In $xpg$, there are two limitations: only for very-well-behaved programs and restriction for $\pginline{Case}$ in Section~\ref{sect:pg}. To be a practical evaluation method, an extension of the target language is needed. After extension of the language, we can evaluate more various programs for experimental results.
Second is introducing an automatic analysis about BX programs and inputs to choose the best evaluation method.
Currently, programmers have to choose the evaluation method by themselves based on their BX programs and inputs. If this analysis is achieved, we can reduce programmers' burden.
Third is introducing a type system to our target language, especially the datatype of sources and view. If we introduce this, we can avoid runtime errors like statically-typed functional languages. For this, we need to investigate more about $construct\_dummy$, because the current definition used in our experiments %Section~\ref{sect:experiments}
will cause type errors.

>>>>

We will continue our work on the following four points.
First is overcoming our limitation in $xpg$. In $xpg$, there are two limitations: only for very-well-behaved programs and restriction for $\pginline{Case}$ in Section~\ref{sect:pg}. To be a practical evaluation method, an extension of the target language is needed. After extension of the language, we can evaluate more various programs for experimental results.
Second is introducing an automatic analysis about BX programs and inputs to choose the best evaluation method.
Currently, programmers have to choose the evaluation method by themselves based on their BX programs and inputs. If this analysis is achieved, we can reduce programmers' burden.
Third is introducing a type system to our target language, especially the datatype of sources and view. If we introduce this, we can avoid runtime errors like statically-typed functional languages. For this, we need to investigate more about how to construct dummy values because the current definition used in our experiments %Section~\ref{sect:experiments}
will cause type errors.
Fourth is using a lazy language. Although we used a strict language OCaml in this paper, if we use a lazy language, we might get laziness for free.
